---
title: "simulation"
author: "Pouya Roudaki"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE,echo=T,include=FALSE}
packages <- c("ks", "quantreg","spatstat","BB","pracma", "akima", "tidyverse", "dplyr", "ggplot2", "parallel", "doParallel","foreach")


#install.packages(packages, dependencies = TRUE)
lapply(packages, require, character.only = TRUE)
```

# Introduction  

This report presents a simulation study comparing the performance of different kernel density estimation approaches against true densities. We evaluate these methods using two metrics: **RKHS MSE** and **MISE**.  

To compute these metrics, we first draw samples from the true density, estimate the densities, and determine the corresponding weights in the kernel exponential family (KEF). We then calculate the **RKHS squared error**:  

$$
SE_\mathcal{H}(\hat{\theta_n}) = ||\hat{\theta_n} - \theta_n||_2^2 = (\hat{\lambda}\hat{\mathbf{w}_n} - \mathbf{w}_n)^{\top} \mathbf{H} \, (\hat{\lambda}\hat{\mathbf{w}_n} - \mathbf{w}_n)
$$  

and the **integrated squared error (ISE)**:  

$$
ISE(\hat{f}) = \int (f - \hat{f})^2 \, dx
$$  

for each sample. This process is repeated **100 times**, allowing us to estimate the **RKHS MSE**:  

$$
MSE_\mathcal{H}(\hat{\theta_n}) = \mathbb{E}[SE_\mathcal{H}(\hat{\theta_n})]=   \mathbb{E}[ (\hat{\lambda}\hat{\mathbf{w}_n} - \mathbf{w}_n)^{\top} \mathbf{H} \, (\hat{\lambda}\hat{\mathbf{w}_n} - \mathbf{w}_n) ]
$$  

and the **Mean Integrated Squared Error (MISE)**:  

$$
MISE(\hat{f}) = \mathbb{E}\left[ISE(\hat{f})\right]  = \mathbb{E}\left[ \int (f - \hat{f})^2 \, dx \right]
$$  

by averaging $ SE_\mathcal{H}(\hat{\theta_n}) $ and $ ISE(\hat{f}) $ across all samples.


# Specify the Density and its domain:

The true density is **mixture normal with 3 peaks**.

```{r}
# Define the weights for the mixture distribution
mixture_weights <- c(1/2,1/6,1/6,1/6)

# Define the parameters for the normal distributions
# First distribution: N(0, 1)
means <- c(0,-1, 0,1)
sds <- c(1,0.1,0.1,0.1)

# Define the domain of density and the corresponding grid
min_x <- -3
max_x <- 3

n_grid <- 10000

grid <-  seq(min_x,max_x,length.out = n_grid)

# Mixture Normal Example
params_mixture <- list(means = means, sds = sds, weights = mixture_weights)
density_characterization_mixture <- list(type = "mixture_normal", parameters = params_mixture)
true_density_grid <- true_density_function(grid, density_characterization_mixture)
```

# Iterative procedure
## Take a Sample

```{r}
n_sample <- 100
sample <- sort(normal_mixture(n_sample, means, sds, mixture_weights))
```

## Find the Cenetered Kernel Matrices

```{r}
centered_kernel_mat_at_sampled <- centered_kernel_matrix(first_vec_kernel = sample,
                                                         second_vec_kernel = sample,
                                                         centering_grid = grid,
                                                         hurst_coef = 0.5)
centered_kernel_mat_at_grid <- centered_kernel_matrix(first_vec_kernel = sample,
                                                      second_vec_kernel = grid,
                                                      centering_grid = grid,
                                                      hurst_coef = 0.5)
centered_kernel_self_grid <- diag(centered_kernel_matrix(first_vec_kernel = grid,
                                                         second_vec_kernel = grid,
                                                         centering_grid = grid,
                                                         hurst_coef = 0.5))
```

## Find the approximated true weights:
```{r}
w_true <- get_true_weights(sample,grid,true_density_mixture)
```

## Estimate the density and kef's weights:
### Fixed Bandwidth Density Estimation 
#### Fixed bandwidth: plug-in
```{r}
start_time <- Sys.time()
estimated_density_grid <- kde(x = sample,h = hpi(sample), eval.points = grid)
end_time <- Sys.time()
# Compute total execution time
pi_time <- difftime(end_time, start_time, units = "secs")
pi_ise <- l2_ise(grid,true_density_grid,estimated_density_grid)
```
#### Fixed bandwidth: Square Cross Validation:
```{r}
start_time <- Sys.time()
estimated_density_grid <- kde(x = sample,h = hscv(sample), eval.points = grid)
end_time <- Sys.time()
# Compute total execution time
scv_time <- difftime(end_time, start_time, units = "secs")
scv_ise <- l2_ise(grid,true_density_grid,estimated_density_grid)
```
#### Fixed bandwidth: Least Square Cross Validation:
```{r}
start_time <- Sys.time()
estimated_density_grid <- kde(x = sample,h = hlscv(sample), eval.points = grid)
end_time <- Sys.time()
# Compute total execution time
lscv_time <- difftime(end_time, start_time, units = "secs")
lscv_ise <- l2_ise(grid,true_density_grid,estimated_density_grid)
```
#### Fixed bandwidth: Normal Scale
```{r}
start_time <- Sys.time()
estimated_density_grid <- kde(x = sample,h = hns(sample), eval.points = grid)
end_time <- Sys.time()
# Compute total execution time
ns_time <- difftime(end_time, start_time, units = "secs")
ns_ise <- l2_ise(grid,true_density_grid,estimated_density_grid)
```
### Kernel Adaptive Density Estimation   
#### Adaptive bandwidth: Adhoc
```{r}
start_time <- Sys.time()
estimated_density_grid <- akj(x = sample, z = grid)
end_time <- Sys.time()
# Compute total execution time
adaptive_adhoc_time <- difftime(end_time, start_time, units = "secs")
adaptive_adhoc_ise <- l2_ise(grid,true_density_grid,estimated_density_grid)
```
#### Adaptive bandwidth: CV optim
```{r}
adaptive_cv_optim <- function(sample, grid, alpha_init, kappa_init, lower_bounds, upper_bounds) {
  
  # Define the objective function to minimize (cv_error)
  objective_function <- function(params) {
    alpha <- params[1]
    kappa <- params[2]
    
    # Adaptive kernel density estimation
    adaptive_result <- akj(sample, grid, alpha = alpha, kappa = kappa)
    
    # Roughness (L2 norm)
    roughness <- pracma::trapz(grid, (adaptive_result$dens)^2)
    
    # Optimized Cross Product Term using proper indexing
    leave_one_out_densities <- vapply(seq_along(sample), function(i) {
      # Remove only the i-th occurrence, not all instances of sample[i]
      sample_wo_i <- sample[-i]
      
      # Compute Adaptive kernel density estimation without the i-th sample point
      adaptive_result_wo_i <- akj(x = sample_wo_i, z = sample[i], alpha = alpha, kappa = kappa)
      
      # Extract density values
      densities_wo_i <- adaptive_result_wo_i$dens
      
      return(densities_wo_i)
    }, numeric(1))  # Ensures output is numeric
    
    # Cross-validation error calculation
    cross_product_sum <- sum(leave_one_out_densities, na.rm = TRUE)
    cv_error <- roughness - (2 / length(sample)) * cross_product_sum
    
    return(cv_error)
  }
  
  # Optimization using Nelder-Mead or BFGS
  result <- optim(
    par = c(alpha_init, kappa_init),   # Initial values
    fn = objective_function,           # Function to minimize
    method = "L-BFGS-B",               # Bounded optimization
    lower = lower_bounds,              # Lower bounds for alpha and kappa
    upper = upper_bounds               # Upper bounds for alpha and kappa
  )
  
  # Extract best parameters
  best_alpha <- result$par[1]
  best_kappa <- result$par[2]
  min_cv_error <- result$value
  
  # Return results
  return(list(best_alpha = best_alpha, best_kappa = best_kappa, min_cv_error = min_cv_error))
}
```

```{r}
start_time_adap <- Sys.time()

alpha_init <- 0.5  # Initial guess for alpha
kappa_init <- 0.5  # Initial guess for kappa
lower_bounds <- c(0.01, 0.01)  # Lower bounds for alpha and kappa
upper_bounds <- c(5, 5)  # Upper bounds for alpha and kappa

# Adaptive hyperparam selection
adaptive_cv_optim <- adaptive_cv_optim(sample, grid, alpha_init, kappa_init, lower_bounds, upper_bounds)

# Adaptive best estimation
estimated_density_grid <- akj(x = sample, z = grid, alpha = adaptive_cv_optim$best_alpha, kappa = adaptive_cv_optim$best_kappa)
end_time_adap <- Sys.time()
adaptive_cv_time <- difftime(end_time_adap, start_time_adap, units = "secs")
adaptive_cv_optim_ise <- l2_ise(grid,true_density_grid,estimated_density_grid)
```
###KEF

#### KEF: Rule of thumb
```{r}
kef_rot <- kef(sample, grid, lambda = 0.1, tau = 0.01/1350)
kef_rot_time <- kef_rot$time 
estimated_density_grid <- kef_rot$probs_grid
kef_rot_ise <- l2_ise(grid,true_density_grid,estimated_density_grid)
kef_rot_se <- rkhs_se(w_hat_vec = kef_rot$weights,
                      w_vec = w_true,
                      kernel_matrix_at_samples = centered_kernel_mat_at_sampled)
```

#### KEF: MLL
```{r}
start_time_mll <- Sys.time()
optimized_mll <- optimize_marginal_log_likelihood_new(centered_kernel_mat_at_sampled,
                                          min_x = min_x,
                                          max_x = max_x,
                                          sample,
                                          initial_lambda = 1,
                                          initial_w = rep(0, length(sampled_x)),
                                          MC_iterations = 10000,
                                          max.iterations = 10,
                                          tol = 1e-3,
                                          parallel_computing = TRUE,
                                          seed = 4)
end_time_mll <- Sys.time()
kef_mll <- kef(sample, grid, lambda = optimized_mll$lambda, tau = optimized_mll$tau)
kef_mll_time <- difftime(as.numeric(kef_mll$time) + as.numeric(difftime(end_time_mll, start_time_mll, units = "secs")),units = "secs")
estimated_density_grid <- kef_mll$probs_grid
kef_mll_ise <- l2_ise(grid,true_density_grid,estimated_density_grid)
kef_mll_se <- rkhs_se(w_hat_vec = kef_mll$weights,
                      w_vec = w_true,
                      kernel_matrix_at_samples = centered_kernel_mat_at_sampled)
```

#### KEF: Jackkinfe optimised MISE
```{r}
kef_jackknife_optim <- function(sample, grid, lambda_init, tau_init, lower_bounds, upper_bounds,loss_metric = "MISE") {
  
  if (loss_metric == "MISE"){
    # Define the objective function to minimize (cv_error)
    objective_function_mise <- function(params) {
      lambda <- params[1]
      tau <- params[2]
      
      # KEF density estimation
      kef_result <- kef(sample, grid, lambda, tau)
      
      # Roughness (L2 norm)
      roughness <- pracma::trapz(grid, (kef_result$probs_grid)^2)
      
      # Optimized Cross Product Term using proper indexing
      leave_one_out_densities <- vapply(seq_along(sample), function(i) {
        # Remove only the i-th occurrence, not all instances of sample[i]
        sample_wo_i <- sample[-i]
        
        # Compute KEF density estimation without the i-th sample point
        kef_result_wo_i <- kef(sample_wo_i, grid, lambda, tau)
        
        # Extract density values
        densities_wo_i <- kef_result_wo_i$probs_sample[i]
        
        return(densities_wo_i)
      }, numeric(1))  # Ensures output is numeric
      
      # Cross-validation error calculation
      cross_product_sum <- sum(leave_one_out_densities, na.rm = TRUE)
      cv_error <- roughness - (2 / length(sample)) * cross_product_sum
      
      return(cv_error)
    }
    
    # Optimization using Nelder-Mead or BFGS
    result <- optim(
      par = c(lambda_init, tau_init),   # Initial values
      fn = objective_function_mise,      # Function to minimize
      method = "L-BFGS-B",               # Bounded optimization
      lower = lower_bounds,              # Lower bounds for alpha and kappa
      upper = upper_bounds               # Upper bounds for alpha and kappa
    )
    
    # Extract best parameters
    best_lambda <- result$par[1]
    best_tau <- result$par[2]
    min_cv_error <- result$value
    
    # Return results
    return(list(best_lambda = best_lambda, best_tau = best_tau, min_cv_error = min_cv_error))
  } else if (loss_metric == "MSE"){
    # Define the objective function to minimize (cv_error)
    objective_function_mse <- function(params) {
      lambda <- params[1]
      tau <- params[2]
      
      # KEF density estimation using whole data
      kef_result <- kef(sample, grid, lambda, tau)
      
      # Estimated weights using full data
      w_estimated_full <- kef_result$weights
      
      # Optimized the Jackknife RKHS Mean Square Error term
      jackknife_error<- mean(vapply(seq_along(sample), function(i) {
        # Remove only the i-th occurrence, not all instances of sample[i]
        sample_wo_i <- sample[-i]
        
        # Compute KEF density estimation without the i-th sample point
        kef_result_wo_i <- kef(sample_wo_i, grid, lambda, tau)
        
        # Extract estimated weights with leave one out data
        w_estimated_wo_i <- kef_result_wo_i$weights
        
        # Estimate the Jackknife error
        jackknife_error_val <- rkhs_se(w_estimated_wo_i, w_estimated_full[-i],     
                                       kernel_matrix_at_samples[-i,-i])
        
        return(jackknife_error_val)
      }, numeric(length(sample))))  # Ensures output is numeric
      
      return(jackknife_error)
    }
    
    # Optimization using Nelder-Mead or BFGS
    result <- optim(
      par = c(lambda_init, tau_init),   # Initial values
      fn = objective_function_mse,       # Function to minimize
      method = "L-BFGS-B",               # Bounded optimization
      lower = lower_bounds,              # Lower bounds for alpha and kappa
      upper = upper_bounds               # Upper bounds for alpha and kappa
    )
    
    # Extract best parameters
    best_lambda <- result$par[1]
    best_tau <- result$par[2]
    min_jackknife_error <- result$value
    
    # Return results
    return(list(best_lambda = best_lambda, best_tau = best_tau, min_jackknife_error = min_jackknife_error))
  } else {
    stop("loss_metric should be `MISE` or `MSE`")
  }
}
```

```{r}
start_time_mise <- Sys.time()

lambda_init <- 0.1  # Initial guess for lambda
tau_init <- 0.01/1350  # Initial guess for tau
lower_bounds <- c(1e-2, 1e-6)  # Lower bounds for lambda and tau
upper_bounds <- c(30, 100)  # Upper bounds for lambda and tau

# kef hyperparam selection
mise_jk_optim <- kef_jackknife_optim(sample, grid, lambda_init, tau_init, lower_bounds, upper_bounds,loss_metric = "MISE")

end_time_mise <- Sys.time()

kef_mise <- kef(sample, grid, lambda = mise_jk_optim$lambda, tau = mise_jk_optim$tau)
kef_mise_jk_time <- difftime(as.numeric(kef_mise$time) + as.numeric(difftime(end_time_mise, start_time_mise, units = "secs")),units = "secs")
estimated_density_grid <- kef_mise$probs_grid
kef_mise_jk_ise <- l2_ise(grid,true_density_grid,estimated_density_grid)
kef_mise_jk_se <- rkhs_se(w_hat_vec = kef_mise$weights,
                      w_vec = w_true,
                      kernel_matrix_at_samples = centered_kernel_mat_at_sampled)
```

#### KEF: Jackkinfe optimised MSE

```{r}
start_time_mse <- Sys.time()

lambda_init <- 0.1  # Initial guess for lambda
tau_init <- 0.01/1350  # Initial guess for tau
lower_bounds <- c(1e-2, 1e-6)  # Lower bounds for lambda and tau
upper_bounds <- c(30, 100)  # Upper bounds for lambda and tau

# kef hyperparam selection
mse_jk_optim <- kef_jackknife_optim(sample, grid, lambda_init, tau_init, lower_bounds, upper_bounds,loss_metric = "MSE")

end_time_mse <- Sys.time()

kef_mse <- kef(sample, grid, lambda = mse_jk_optim$lambda, tau = mse_jk_optim$tau)
kef_mse_jk_time <- difftime(as.numeric(kef_mse$time) + as.numeric(difftime(end_time_mse, start_time_mse, units = "secs")),units = "secs")
estimated_density_grid <- kef_mse$probs_grid
kef_mse_jk_ise <- l2_ise(grid,true_density_grid,estimated_density_grid)
kef_mse_jk_se <- rkhs_se(w_hat_vec = kef_mse$weights,
                      w_vec = w_true,
                      kernel_matrix_at_samples = centered_kernel_mat_at_sampled)
```

## Final Result

Store the final result here.
```{r}
result_iter_i <- data.frame(matrix(data = NA, nrow = 10, ncol = 4))
colnames(result_iter_i) <- c("method","time","ISE","RKHS_SE")
result_iter_i$method <- c("fixed_pi", "fixed_scv","fixed_lscv","fixed_ns",
                          "adaptive_adhoc","adaptive_cv_optim",
                          "kef_rot","kef_mll", "kef_jk_mise_optim", "kef_jk_mse_optim")

result_iter_i$time <- c(pi_time,scv_time,lscv_time,ns_time,adaptive_adhoc_time,
                        adaptive_cv_time,kef_rot_time,kef_mll_time,
                        kef_mise_jk_time,kef_mse_jk_time)

result_iter_i$ISE <- c(pi_ise,scv_ise,lscv_ise,ns_ise,adaptive_adhoc_ise,
                        adaptive_cv_optim_ise,kef_rot_ise,kef_mll_ise,
                        kef_mise_jk_ise,kef_mse_jk_ise)

result_iter_i$SE <- c(NA,NA,NA,NA,NA,
                      NA,kef_rot_se,kef_mll_se,
                      kef_mise_jk_se,kef_mse_jk_se)


```



